---
title: "Mid-course project"
subtitle: "EC 421: Introduction to Econometrics"
# author: "Edward Rubin"
date: "Due *before* midnight on Tuesday, 12 May 2020"
# date: ".it.biggest[Solutions]"
output:
  xaringan::moon_reader:
    css: ['default', 'metropolis', 'metropolis-fonts', 'my-css.css']
    # self_contained: true
    nature:
      ratio: '8.5:11'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
layout: true
class: clear
---

```{r, setup, include = F}
# Knitr options
library(knitr)
opts_chunk$set(
  comment = "#>",
  fig.align = "center",
  fig.height = 7,
  fig.width = 10.5,
  warning = F,
  message = F
)
opts_chunk$set(dev = "svg")
options(device = function(file, width, height) {
  svg(tempfile(), width = width, height = height)
})
options(digits = 4)
options(width = 90)
```

## Instructions

.mono.b[DUE:] **One member** of your group must upload your answer on [Canvas](https://canvas.uoregon.edu/) *before* midnight on Tuesday, 12 May 2020. All members of the group must be listed on the submission.

.mono.b[IMPORTANT:] As with your homework, you must submit .b[two files]:
<br> .b.mono[1.] your typed responses/answers to the question (in a Word file or something similar)
<br> .b.mono[2.] the .mono[R] script you used to generate your answers. Each student must turn in her/his own answers.

If you are using RMarkdown, you can submit a single file.

.mono.b[README!] The last page has a table that describes each variable in the dataset (`proj1.csv`).

.mono.b[INTEGRITY:] Groups can either have **one or two members**. Only one person needs to submit your final document. If you are suspected of cheating in any way (for example, copying from someone else), then you will receive a zero. We may report you to the dean.

.mono.b[GRADING:] Your grade for this project will be based upon the accuracy of your answers *and* how well you explain/illustrate your answers. We value short, accurate answers over long, meandering answers. Edit your answers!

.mono.b[EMAIL POLICY:] Do not ask the GEs or the instructor for help coding or for help answering these questions. You may only ask **clarifying** questions. Use Google and the course's materials (lectures, labs, notes, assignment keys).

---

## Questions

.b[01.] Summarize and describe the dataset. Your answer should include:

- What share of the sample is female? What share of the sample is non-white?
- How skewed are the income and education distributions?
- Create three figures (graphs) that summarize key variables.
- Create two figures (graphs) that summarize how key variables relate to each other.

Explain your decisions on summarizing the data. What do you learn about potential relationships?

.b[02.] Regress individuals' income (`income`) on an intercept and their education (`education`).

.b[03.] Create a scatter plot with the residuals from .b[02] on the y axis and education on the x axis.

.b[04.] Does the scatter plot from .b[03] suggest that **heteroskedasticity** may be present? Explain your answer.

.b[05.] More generally: Does the scatter plot from .b[03] suggest that there are any issues with **your specification**? Explain.

.b[06.] Explain why the regression in .b[02] could suffer from omitted-variable bias. 

.b[07.] Give an example of an omitted variable (other than *ability*) that could cause bias in the regression in .b[02]. Just to be clear: Do not use the variable *ability* as your example. 

- Explain how your example variable satisfies both requirements for omitted-variable bias.
- Describe the direction of the bias this variable would cause (when we estimate the effect of education on income). Explain your answer.

.b[08.] Now regress *income* on an intercept, *education*, **and** *ability*. Interpret the results.

.b[09.] Does your estimate for the effect of education on income change from question .b[02] to question .b[08]? Explain why this change (or lack of change) makes sense. 

*Hint:* Is there a significant relationship between *education* and *ability*?

.b[10.] Up to this point, we have generally told you which regressions to run. And we've stuck with pretty simple regressions (*e.g.*, regress `y` on `x1` and `x2`). We now want you to explore the actual complexity of econometric/statistical analyses. Estimate three new models. These models should not match your previous models (in .b[02] and .b[08]). Across these three new models, you should include (at least once):

- a log-transformed variable (*i.e.*, use `log`)
- an interaction 

.b[11.] How did you choose your specifications in .b[10]? Explain your decision making.

.b[12.] Which of your new models is "best"—if you must choose one model, which would you choose? Why?

.b[13.] For your "best" model (chosen in .b[12]): Interpret the coefficients and comment on their statistical significance.

.b[14.] Do you *trust* the estimates from your *best model*? Explain why/why not.

.b[15.] Suppose you want to estimate the effect of high-school graduation. How could you use the current data to estimate this effect? Describe any regressions, estimates, figures, and/or caveats you would make. *Note:* You can assume that someone with 12 years of education graduated from high school.

---

```{r, background variables, echo = F, message = F}
# Load requisite packages
pacman::p_load(tidyverse, knitr, kableExtra, here)
# Load data
proj_df = here("proj1.csv") %>% read_csv()
# Create table of variable descriptions
var_tbl = data.frame(
  Variable = names(proj_df) %>% paste0(".mono-small[", ., "]"),
  Description = c(
  	"The individual's annual income (in US dollars).",
  	"Binary indicator (1,0) for whether the individual is currently married.",
  	"Number of children.",
  	"Binary indicator (1,0) for whether the individual belongs to a non-white ethnicity.",
  	"Binary indicator (1,0) for whether the individual is female.",
  	"Years of education.",
  	"Binary indicator (1,0) for whether the individual's home is classified as 'urban'.",
  	"Ability on scale of 0 (lowest) to 100 (highest)."
  )
)
kable(var_tbl) %>%
  kable_styling(full_width = F)
```

---
exclude: true

```{r, print pdf, echo = F, eval = F}
pagedown::chrome_print("prompt.html", output = "project1.pdf")
```